---
title: "optim_stat"
author: "Andrew Martin"
date: "3/27/2018"
output: html_document
---

source `constants.R` and `data.R` functions to get proj system data into memory:
```{r}
source(file = 'constants.R')
source(file = 'data.R')

```

```{r}
library(tidyverse)

```

We want to re-run optim2.Rmd, but instead of aggregating the zscores, include the individual zscores per player.

```{r}

player_univ <- all_proj$steam$h_final %>% 
  arrange(-final_zsum) %>% 
  dplyr::top_n(10, wt = final_zsum)
  
num_players <- nrow(player_univ)
roster_names <- c("C", "1B", "2B")
num_roster <- 3
num_stat <- 5

best <- make.lp(0,  num_players * num_roster * num_stat)

best

```

We need to define the variables to make this model interpretable.

We'll use the order
player-stat-position

```{r}

player_vars <- map(
  player_univ$mlbid, 
  function(.x) rep(.x, num_stat * num_roster)
) %>% 
reduce(c)

stat_vars <- rep(all_proj$steam$h_final[9:13] %>% names(), num_players * num_roster)

pos_base <- map(
  seq(1:num_roster), 
  function(.x) rep(.x, num_stat)
) %>% 
reduce(c)

pos_vars <- rep(pos_base, num_players)

player_vars
stat_vars
pos_vars

full_vars <- paste(player_vars, stat_vars, pos_vars, sep = '_')
full_vars

```

Add constraints

First add the player constraints.
No player can be in more than one position
```{r}



```


