---
title: "optim_stat"
author: "Andrew Martin"
date: "3/27/2018"
output: html_document
---

source `constants.R` and `data.R` functions to get proj system data into memory:
```{r}
source(file = 'constants.R')
source(file = 'data.R')

```

```{r}
library(tidyverse)
library(lpSolveAPI)

```

We want to re-run optim.Rmd, but instead of aggregating the zscores, include the individual zscores per player.

```{r}

player_univ <- all_proj$steam$h_final %>% 
  arrange(-final_zsum) %>% 
  dplyr::top_n(10, wt = final_zsum)
  
num_players <- nrow(player_univ)
roster_names <- c("C", "1B", "2B")
num_roster <- 3

#best <- make.lp(0,  (num_players * num_roster) + num_stat)
best <- make.lp(0,  num_players * num_roster)

best

```

We need to define the variables to make this model interpretable.

We'll use the order
player-stat-position

```{r}

player_vars <- map(
  player_univ$mlbid, 
  function(.x) rep(.x, num_roster)
) %>% 
reduce(c)

pos_vars <- rep(roster_names, num_players)

player_vars
pos_vars

full_vars <- paste(player_vars, pos_vars, sep = '_')
price_vars <- paste0(player_univ$mlbid, '_price')

final_vars <- c(full_vars, price_vars)
final_vars
```

Add constraints

First add the player constraints.
No player can be in more than one position.

```{r}

for (i in seq(1:num_players)) {
  this_id <- player_univ$mlbid[i]
  this_player <- player_univ[player_univ$mlbid == this_id, ]$fullname
  print(this_player)
  
  add.constraint(
    lprec = best, 
    xt = rep(1, num_roster),
    type = "<=",
    rhs = 1,
    indices = which(player_vars %in% this_id)
  )
}

best
```

Next add roster constraints.
No position can have more than one player.

```{r}

for (i in seq(1:num_roster)) {
  this_pos <- roster_names[i]
  print(this_pos)
  
  add.constraint(
    lprec = best, 
    xt = rep(1, num_players),
    type = "<=",
    rhs = 1,
    indices = which(pos_vars %in% this_pos)
  )
}

best

```

The objective function will be the value of the players at each of the positions

```{r}

player_value <- player_univ$unadjusted_zsum

player_value <- map(
  player_value, 
  function(.x) rep(.x, num_roster)
) %>% 
reduce(c)

set.objfn(best, player_value)

best


```

Pad the objective function with zero for price dummies

```{r}
#pad with zeros for price dummies
player_value

```


We want to maximize value

```{r}
lp.control(best, sense="max")

best
```

```{r}

solve(best)
get.objective(best)
get.variables(best)
```

