---
title: "optim"
author: "Andrew Martin"
date: "3/22/2018"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}

library(stats) # normally loaded

eurodistmat <- as.matrix(eurodist)

distance <- function(sq) {  # Target function
    sq2 <- embed(sq, 2)
    sum(eurodistmat[cbind(sq2[,2], sq2[,1])])
}

genseq <- function(sq) {  # Generate new candidate sequence
    idx <- seq(2, NROW(eurodistmat)-1)
    changepoints <- sample(idx, size = 2, replace = FALSE)
    tmp <- sq[changepoints[1]]
    sq[changepoints[1]] <- sq[changepoints[2]]
    sq[changepoints[2]] <- tmp
    sq
}

sq <- c(1:nrow(eurodistmat), 1)  # Initial sequence: alphabetic
distance(sq)
# rotate for conventional orientation
loc <- -cmdscale(eurodist, add = TRUE)$points
x <- loc[,1]; y <- loc[,2]
s <- seq_len(nrow(eurodistmat))
tspinit <- loc[sq,]

plot(x, y, type = "n", asp = 1, xlab = "", ylab = "",
     main = "initial solution of traveling salesman problem", axes = FALSE)
arrows(tspinit[s,1], tspinit[s,2], tspinit[s+1,1], tspinit[s+1,2],
       angle = 10, col = "green")
text(x, y, labels(eurodist), cex = 0.8)

set.seed(123) # chosen to get a good soln relatively quickly

res <- optim(
  par = sq, 
  fn = distance, 
  gr = genseq, 
  method = "SANN",
  control = list(maxit = 30000, temp = 2000, trace = TRUE, REPORT = 500)
)
res  # Near optimum distance around 12842


```

source the `data.R` functions to get proj system data into memory:
```{r}
source(file = 'constants.R')
source(file = 'data.R')

```


```{r}


steamer <- all_proj$steam$h_final
steamer$final_zsum <- steamer$final_zsum * -1

steamer$rn <- row.names(steamer)
#my_team <- c(454560, 276055, 457787, 488703, 502003)
my_team <- c(1, 10, 169, 314, 576)

# steamer %>%
#   filter(mlbid %in% my_team) %>%
#   print.AsIs()

names(my_team) <- c('C', '1B', '2B', 'SS', '3B' )

eval_team <- function(team_vector) {
  #get matching ids
  this_team <- steamer %>%
    filter(rn %in% team_vector) 
  
  #how to score?
  score <- this_team %>% pull(final_zsum)
  
  sum(score) 
}

change_team <- function(tv) {
  #pick a position to change
  random_player <-  sample(tv, size = 1)
  to_change <- names(random_player)

  #players matching the position
  pos_pool <- steamer %>%
    filter(priority_pos == to_change) %>%
    pull(rn)

  #replace the position to change
  new_player <- sample(pos_pool, 1)
  tv[to_change] <- new_player
  
  as.numeric(tv)
}

best <- optim(
  par = my_team,
  fn = eval_team,
  gr = change_team,
  #control = list(fnscale = -10, trace = 1, maxit = 1000)
  method = "SANN",
  control = list(maxit = 30000, temp = 2000, trace = TRUE, REPORT = 500)
)

best
best$par

best$message

steamer %>%
  filter(rn %in% best$par) %>%
  select(rn, fullname, priority_pos, ab, r, rbi, sb, tb, obp) %>%
  print.AsIs()

```

that's the wrong optimization approach above.  this is actually a mixed-integer linear programming problem

```{r}

library(lpSolveAPI)

```

lets imagine 3 players across 2 roster spots
```{r}

num_players <- 3
num_roster <- 2

```

the constraints are as follows:
- each player can only occupy one roster spot (1 constraint per player)
- each roster spot can only have one player (1 constraint per roster)

```{r}

player_appears_once <- num_players
roster_used_once <- num_roster

num_constraints <- player_appears_once + roster_used_once

best <- make.lp(0,  num_players * num_roster)

best
```

I interpret the 6 variables to be a combination of each player/roster pair.
First let's add the constraint on each player.  In this reading, C1 and C2 relate to player 1, C3 and C4 to player 2, C5 and C6 to player 3.

```{r}

add.constraint(best, c(1, 1, 0, 0, 0, 0), type = '<=', rhs = 1)
add.constraint(best, c(0, 0, 1, 1, 0, 0), type = '<=', rhs = 1)
add.constraint(best, c(0, 0, 0, 0, 1, 1), type = '<=', rhs = 1)

best
```

Then let's add the constraint on each roster spot. 
In interpret C1:C3 as relating to roster 1.  C4:C6 relate to roster spot 2.

```{r}

add.constraint(best, c(1, 0, 1, 0, 1, 0), type = '=', rhs = 1)
add.constraint(best, c(0, 1, 0, 1, 0, 1), type = '=', rhs = 1)

best
```

We know that in this interpretation the values can only be 0 or 1.  So we can set the upper bound to 1.

```{r}

set.bounds(best, upper = c(1, 1, 1, 1, 1, 1))

best
```

The objective function will be the value of the players at each of the positions

```{r}
all_proj$steam$h_final %>% arrange(-final_zsum) %>% head() %>% select(fullname, priority_pos, unadjusted_zsum)
#trout, altuve, goldschmidt

set.objfn(best, c(11,11,8.7,8.7,8.19,8.19))

best

```

We want to maximize value

```{r}
lp.control(best, sense="max")

best
```

```{r}

solve(best)
get.objective(best)
get.variables(best)
```

Ok, so we have an optimal solution!
Recall that our variables map to players.  C1:C2 is player 1, C3:C4 is player 2, C5:C6 is player 3.
So we interpret that as:
Optimal solution is 
-player 1 and player 2
-for player 1, roster slot 1.
-for player 2, roster slot 2.

That matches what we think should happen!


