---
title: "optim_stat"
author: "Andrew Martin"
date: "3/27/2018"
output: html_document
---

source `constants.R` and `data.R` functions to get proj system data into memory:
```{r}
source(file = 'constants.R')
source(file = 'data.R')

```

```{r}
library(tidyverse)
library(lpSolveAPI)

```

```{r}

print.lpExtPtr <- function (x, ...) 
{
    m <- dim(x)[1]
    n <- dim(x)[2]
    control <- lp.control(x)
    if (n < 1) {
        cat(paste("Model name: ", name.lp(x), "\n", sep = ""))
        return(invisible(x))
    }
    if (n > 800) {
        cat(paste("Model name: ", name.lp(x), "\n", "  a linear program with ", 
            n, " decision variables and ", m, " constraints\n", 
            sep = ""))
        return(invisible(x))
    }
    ans <- matrix(0, m + 1, n)
    for (j in 1:n) {
        col <- get.column(x, j)
        ans[1 + col$nzrow, j] <- col$column
    }
    type <- get.type(x)
    type[type == "integer"] <- "Int"
    type[type == "real"] <- "Real"
    kind <- get.kind(x)
    kind[kind == "standard"] <- "Std"
    kind[kind == "semi-continuous"] <- "S-C"
    bounds <- get.bounds(x)
    upper <- bounds$upper
    lower <- bounds$lower
    ans <- format(rbind(dimnames(x)[[2]], ans, kind, type, upper, 
        lower), justify = "right")
    sense <- ifelse(control$sense == "minimize", "Minimize", 
        "Maximize")
    lhs <- get.constr.value(x, side = "lhs")
    rhs <- get.constr.value(x, side = "rhs")
    rowNames <- format(c("", sense, dimnames(x)[[1]], "Kind", 
        "Type", "Upper", "Lower"))
    constrs <- format(c("", "", get.constr.type(x), "", "", "", 
        ""), justify = "right")
    rhs <- format(c("", "", as.character(rhs), "", "", "", ""), 
        justify = "right")
    print.lhs <- any(!is.infinite(lhs[is.element(get.constr.type(x, 
        as.char = FALSE), c(1, 2))]))
    lhs <- format(c("", "", as.character(lhs), "", "", "", ""), 
        justify = "right")
    if (print.lhs) 
        ans <- cbind(rowNames, lhs, constrs, ans, constrs, rhs)
    else ans <- cbind(rowNames, ans, constrs, rhs)
    ans <- apply(ans, 1, paste, collapse = "  ")
    ans <- paste(ans, collapse = "\n")
    model.name <- paste("Model name: ", name.lp(x), "\n", sep = "")
    ans <- paste(model.name, ans, "\n", sep = "")
    cat(ans)
    invisible(x)
}
```

We want to re-run optim.Rmd, but instead of aggregating the zscores, include the individual zscores per player.

```{r}

player_univ <- all_proj$steam$h_final %>% 
  arrange(-final_zsum) %>% 
  dplyr::top_n(2, wt = final_zsum)
  
num_players <- nrow(player_univ)
roster_names <- c("C", "1B", "2B", "SS", "3B", "OF")
num_roster <- 6

best <- make.lp(0,  num_players * num_roster)

best

```

We need to define the variables to make this model interpretable.

We'll use the order
player-stat-position

```{r}

player_vars <- map(
  player_univ$mlbid, 
  function(.x) rep(.x, num_roster)
) %>% 
reduce(c)

pos_vars <- rep(roster_names, num_players)

player_vars
pos_vars

full_vars <- paste(player_vars, pos_vars, sep = '_')

final_vars <- full_vars
final_vars
```

Add constraints

First add the player constraints.
No player can be in more than one position.
Players can only play at positions for which they are eligibile.

```{r}

for (i in seq(1:num_players)) {
  this_id <- player_univ$mlbid[i]
  player_df <- player_univ[player_univ$mlbid == this_id, ]
  this_player <- player_df$fullname
  print(this_player)
  
  #which positions can they play?
  player_indices <- which(player_vars %in% this_id)
  player_pos <- strsplit(player_df$position, ', ') %>% reduce(c)
  position_indices <- roster_names %in% player_pos
  eligib_positions <- roster_names[position_indices]
  
  add.constraint(
    lprec = best, 
    xt = rep(1, length(eligib_positions)),
    type = "<=",
    rhs = 1,
    indices = player_indices[position_indices]
  )
}

best
write.lp(best, "best.lp", "lp") 
```

Next add roster constraints.
No position can have more than one player.

```{r}

for (i in seq(1:num_roster)) {
  this_pos <- roster_names[i]
  print(this_pos)
  
  add.constraint(
    lprec = best, 
    xt = rep(1, num_players),
    type = "<=",
    rhs = 1,
    indices = which(pos_vars %in% this_pos)
  )
}

best

```

The objective function will be the value of the players at each of the positions

```{r}

player_value <- player_univ$unadjusted_zsum

player_value <- map(
  player_value, 
  function(.x) rep(.x, num_roster)
) %>% 
reduce(c)

set.objfn(best, player_value)

best


```

We want to maximize value

```{r}
lp.control(best, sense="max")

best
```

```{r}

solve(best)
total_value <- get.objective(best)

optimal <- get.variables(best)

total_value
optimal
```

Interpret the result

```{r}

optimal_players <- final_vars[optimal == 1]
optimal_players

```
add in names

```{r}

#gsub('(.)_(.)', optimal_players, value = TRUE)

just_players <- map_chr(
  strsplit(optimal_players, '_'), 
  function(.x) .x %>% extract2(1)
)

just_pos <- map_chr(
  strsplit(optimal_players, '_'), 
  function(.x) .x %>% extract2(2)
)

player_univ[player_univ$mlbid %in% just_players, ]$fullname

```
