---
title: "optim_stat"
author: "Andrew Martin"
date: "3/27/2018"
output: html_document
editor_options: 
  chunk_output_type: console
---

source `constants.R` and `data.R` functions to get proj system data into memory:

```{r}

source(file = 'constants.R')
source(file = 'data.R')

```

```{r}

library(tidyverse)
library(lpSolveAPI)

```

Building on optim4 - adding in the second-layer player constraints for multi-elig players.

Instead of generating the model right up front, like I did in 2 and 3, do it later, after we know position eligibility.

```{r}

player_univ <- all_proj$steam$h_final %>% 
  arrange(-final_zsum) %>% 
  dplyr::top_n(50, wt = final_zsum)
  
num_players <- nrow(player_univ)
# roster_names <- c("C", "1B", "2B", "SS", "3B", "OF", "OF", "OF")
# num_roster <- 8

roster_names <- c("C", "1B", "2B", "SS", "3B", "OF")
num_roster <- 6

num_stat <- 5
stat_names <- c('r', 'rbi', 'sb', 'tb', 'obp')

multipos_players <- player_univ %>%
  filter(grepl(',', position, fixed = TRUE))

multipos_players
```

We need to define the variables to make this model interpretable.

We'll use the order
player-position-stat

```{r}

player_only <- list()
pos_only <- list()
player_pos <- list()
player_pos_stat <- list()

for (i in 1:nrow(player_univ)) {
  this_row <- player_univ[i, ]
  this_pos <- strsplit(this_row$position, ', ')[[1]]
  this_player_pos <- paste(this_row$mlbid, this_pos, sep = '_')
  this_player_pos_stat <- map(
    this_player_pos, 
    function(.x) paste(.x, stat_names, sep = '_')
  ) %>% reduce(c)
  
  player_only[[i]] <- rep(this_row$mlbid, length(this_player_pos_stat))
  pos_only[[i]] <- gsub(
    "(\\d+)([[:punct:]]+)([[:alnum:]]+)(.+)", 
    "\\3", 
    this_player_pos_stat
  )
  player_pos[[i]] <- gsub(
    "(\\d+)([[:punct:]]+)([[:alnum:]]+)(.+)", 
    "\\1\\2\\3", 
    this_player_pos_stat
  )
  player_pos_stat[[i]] <- this_player_pos_stat

}

player_only <- player_only %>% reduce(c)
pos_only <- pos_only %>% reduce(c)
player_pos <- player_pos %>% reduce(c)
player_pos_stat <- player_pos_stat %>% reduce(c)
player_pos
player_pos_stat
```


The size of the model will be:
- one decision variable for every position a player under consideration is _eligible_ for, for _every stat_.  If a player has multi-position eligibility we simply create as many decision variables as positions.

- one decision variable for every player that contains the expected _price_ of the player (using an ADP model)

- one decision variable for every player that contains the price paid of the player

```{r}

unq_player_pos <- player_pos %>% unique()
d_team_dummy <- paste0(unq_player_pos, '_discontin')
d_multipos_dummy <- paste0(multipos_players$mlbid, '_discontin')
d_pos_dummy <- paste0(roster_names, '_discontin')
expected_price <- paste0(player_univ$mlbid, '_exp_price')
auction_price <- paste0(player_univ$mlbid, 'auction_price')
#final_vars <- c(player_pos, d_team_dummy, expected_price, auction_price)
final_vars <- c(player_pos, d_team_dummy, d_multipos_dummy, d_pos_dummy)

best <- make.lp(0,  length(final_vars), verbose = 'detailed')

best

```

Add constraints

First add the player constraints.
Players can only play at positions for which they are eligibile (enforced by how we created our decision variables above).

Players must play one position.  That means that the 5 stats need to travel as a group.  We create this as a discontinuous variable in lpsolve.

THIS WAS HARD
cribbing off of
https://stackoverflow.com/questions/40900940/linear-programming-with-conditional-constraints-in-r

I believe if I create a constraint that has:
- 1s in the columns that represent the player's 5 stats
- a -1 in the column that is the dummy decision variable that tracks the 5 stats
- a lower bound of 5 (or 4.9 from the example?)

it will force the 5 stats to travel together.  somehow.

Update: This needs to iterate over every _player position_ because you can only play, say, Kyle Schwarber at C *or* at OF.  You can't play a smattering of both C and OF.
We will _also_ need to create a second layer constraint for multi-position eligibility players, so that only one Kyle Schwarber is used by the algorithm.  This stuff is subtly complicated!

```{r}

for (i in seq(1:length(unq_player_pos))) {
  this_id <- gsub("(\\d+)([[:punct:]]+)([[:alnum:]]+)", "\\1", unq_player_pos[i])
  this_player_pos <- unq_player_pos[i]
  player_df <- player_univ[player_univ$mlbid == this_id, ]
  this_player <- player_df$fullname
  print(this_player)
  
  #find player indices
  player_indices <- which(player_pos %in% this_player_pos)
  #find player discontinuity dummy
  p_discontin_index <- which(final_vars == paste0(this_player_pos, '_discontin'))

  add.constraint(
    lprec = best, 
    xt = c(rep(1, length(player_indices)), -1),
    type = ">=",
    rhs = 5,
    indices = c(player_indices, p_discontin_index)
  ) 
}

best
```

SECOND LAYER CONSTRAINTS FOR PLAYERS WITH MULTIELIG 

```{r}

for (i in 1:nrow(multipos_players)) {
  player_df <- multipos_players[i, ]
  print(player_df$fullname)
  
  #find player indices
  player_indices <- which(player_only %in% player_df$mlbid)
  #find player discontinuity dummy
  p_discontin_index <- which(final_vars == paste0(player_df$mlbid, '_discontin'))

  add.constraint(
    lprec = best, 
    xt = c(rep(1, length(player_indices)), -1),
    type = ">=",
    rhs = 5,
    indices = c(player_indices, p_discontin_index)
  )
}

best

```

ADD POSITION CONSTRAINTS

```{r}

for (i in 1:length(roster_names)) {
  this_roster <- roster_names[i]
  print(this_roster)
  
  #find position indices
  pos_indices <- which(pos_only %in% this_roster)
  #find player discontinuity dummy
  pos_discon_index <- which(final_vars == paste0(this_roster, '_discontin'))
  all_pos_indices <- c(pos_indices, pos_discon_index)
  print(all_pos_indices)
  
  add.constraint(
    lprec = best, 
    xt = c(rep(1, length(pos_indices)), -1),
    type = ">=",
    rhs = 5,
    indices = all_pos_indices
  )
}

best
```


Next add price constraints

```{r}


```

Next add expected price constraints

```{r}


```


The objective function will be the value of the players at each of the positions plus dummies for price/expected price.

```{r}

player_stat_zscore <- function(playerid, stat) {
  df <- player_univ %>%
    filter(mlbid == playerid)
  
  df %>% extract2(paste0(stat, '_zscore'))
}

player_stat_value <- map(
  player_pos_stat,
  function(.x) {
    player <- gsub("(\\d+)([[:punct:]]+)([[:alnum:]]+)(.+)", "\\1", .x)
    stat <- gsub("(\\d+)([[:punct:]]+)([[:alnum:]]+)([[:punct:]]+)(.+)", "\\5", .x)
    player_stat_zscore(player, stat)
  }
) %>%
reduce(c)

```

For price and expected price, the objective function is zero, because we aren't trying to maximize on those - they simply constrain our solution.

Add zeros:

```{r}

# final_obj <- c(
#   player_stat_value, 
#   rep(0, length(expected_price)),
#   rep(0, length(auction_price))
# )

final_obj <- c(
  player_stat_value,
  #all the various discon constraints
  rep(0, length(final_vars) - length(player_stat_value))
)

```

Set the objective function

```{r}

set.objfn(best, final_obj)

best

```


Player stats are either yes/no (1/0) so we set to binary.
We want to maximize value

Set the bounds for all discon variables


```{r}

all_discon_indices <- which(grepl('_discon', final_vars))
set.semicont(best, col = all_discon_indices)
set.bounds(best, upper = rep(5, length(all_discon_indices)), col = all_discon_indices)

```


```{r}

set.type(best, col = seq(1:length(player_stat_value)), type = "binary") 

lp.control(best, sense="max")

best
write.lp(best, "best.lp", "lp") 
write.lp(best, "best.mps", "mps")
```

```{r}

solve(best)
total_value <- get.objective(best)

optimal <- get.variables(best)

total_value
optimal
```

Interpret the result

```{r}

pos_only[optimal == 1] %>% table()

player_only[optimal == 1] %>% table()

```


```{r eval = FALSE}


optimal_players <- player_pos_stat[optimal == 1]
optimal_players

```

add in names

```{r eval = FALSE}

#gsub('(.)_(.)', optimal_players, value = TRUE)

just_players <- map_chr(
  strsplit(optimal_players, '_'), 
  function(.x) .x %>% extract2(1)
)

just_pos <- map_chr(
  strsplit(optimal_players, '_'), 
  function(.x) .x %>% extract2(2)
)

player_univ[player_univ$mlbid %in% just_players, ]$fullname

```
